---
title: "UCSB Department of Recreation"
subtitle: "Instagram Analytics — Spring Quarter 2026"
author: "UCSB Rec Marketing"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    number_sections: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,   # hide code in output — set to TRUE if you want to see it
  message = FALSE,
  warning = FALSE,
  fig.width  = 10,
  fig.height = 5.5,
  dpi = 150
)
```

```{r packages}
required_packages <- c("tidyverse", "lubridate", "scales",
                        "patchwork", "knitr", "kableExtra", "ggtext")
to_install <- required_packages[!required_packages %in% rownames(installed.packages())]
if (length(to_install)) install.packages(to_install, quiet = TRUE)

library(tidyverse)
library(lubridate)
library(scales)
library(patchwork)
library(knitr)
library(kableExtra)
library(ggtext)
```

```{r theme}
# ── UCSB brand colors ──────────────────────────────────────────────────────────
ucsb_blue  <- "#003660"
ucsb_gold  <- "#FEBC11"
ucsb_navy  <- "#002542"
ucsb_light <- "#9CBEBE"
ucsb_grey  <- "#BBC8CB"

theme_ucsb <- function() {
  theme_minimal(base_size = 12) +
    theme(
      plot.title       = element_text(color = ucsb_blue, face = "bold", size = 14),
      plot.subtitle    = element_text(color = "grey40",  size = 10),
      plot.caption     = element_text(color = "grey55",  size = 8),
      axis.title       = element_text(color = ucsb_navy, size = 10),
      axis.text        = element_text(color = "grey30"),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "grey92"),
      legend.position  = "bottom",
      strip.text       = element_text(color = ucsb_blue, face = "bold")
    )
}
```

```{r load-data}
# ── Load cleaned CSVs ──────────────────────────────────────────────────────────
if (rstudioapi::isAvailable()) {
  script_dir <- dirname(rstudioapi::getSourceEditorContext()$path)
  if (nchar(script_dir) > 0) setwd(script_dir)
}

needed <- c("spring_posts_clean.csv", "spring_reels_clean.csv", "spring_growth_clean.csv")
missing <- needed[!file.exists(needed)]
if (length(missing)) stop("Missing files: ", paste(missing, collapse = ", "))

day_levels <- c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")

posts <- read_csv("spring_posts_clean.csv", show_col_types = FALSE) |>
  mutate(
    Time_Posted = ymd_hms(Time_Posted),
    Date        = as.Date(Date),
    Media_Type  = str_to_title(Media_Type),
    DayOfWeek   = factor(DayOfWeek, levels = day_levels),
    Month       = factor(Month, levels = c("April","May","June"))
  )

reels <- read_csv("spring_reels_clean.csv", show_col_types = FALSE) |>
  mutate(
    Time_Posted = ymd_hms(Time_Posted),
    Date        = as.Date(Date),
    Media_Type  = str_to_title(Media_Type),
    DayOfWeek   = factor(DayOfWeek, levels = day_levels),
    Month       = factor(Month, levels = c("April","May","June"))
  )

growth <- read_csv("spring_growth_clean.csv", show_col_types = FALSE) |>
  mutate(
    Date      = as.Date(Date),
    DayOfWeek = factor(DayOfWeek, levels = day_levels)
  )
```

---

# Executive Summary

Spring Quarter 2026 ran from **April 6 – May 18, 2026** (data through end of quarter export).
UCSB Recreation published **`r nrow(posts)` posts** and **`r nrow(reels)` reels**, achieving
an average engagement rate of **`r round(mean(posts$Engagement_Rate_Pct), 1)`%** on posts —
up from 3.65% in Winter Quarter. No image-only posts were published this quarter;
all content was either **carousel** (`r sum(posts$Media_Type == "Carousel")` posts) or
**video** (`r sum(posts$Media_Type == "Video")` posts).

A new data source is available this quarter: **daily Profile Growth & Discovery** tracking,
which captures follower counts, daily views, and daily reach — enabling trend analysis
not possible with Winter data.

---

# Profile Growth & Discovery {.tabset}

## Follower Trend

```{r follower-trend, fig.height=4}
growth_filtered <- growth |> filter(!is.na(Followers))

ggplot(growth_filtered, aes(x = Date, y = Followers)) +
  geom_line(color = ucsb_blue, linewidth = 1.1) +
  geom_point(color = ucsb_gold, size = 2.5) +
  geom_smooth(method = "lm", se = FALSE, color = ucsb_light,
              linetype = "dashed", linewidth = 0.8) +
  scale_y_continuous(labels = comma, limits = c(NA, NA)) +
  scale_x_date(date_breaks = "1 week", date_labels = "%b %d") +
  labs(
    title    = "Follower Count — Spring Quarter 2026",
    subtitle = "Dashed line = linear trend",
    x = NULL, y = "Followers",
    caption  = "Source: UCSB Rec Instagram Profile Growth data"
  ) +
  theme_ucsb()
```

```{r follower-stats}
follower_start <- growth_filtered |> filter(Date == min(Date)) |> pull(Followers)
follower_end   <- growth_filtered |> filter(Date == max(Date)) |> pull(Followers)
net_growth     <- follower_end - follower_start

tibble(
  Metric  = c("Followers at Start", "Followers at End", "Net Growth", "Growth Rate"),
  Value   = c(
    comma(follower_start),
    comma(follower_end),
    paste0("+", net_growth),
    paste0(round(net_growth / follower_start * 100, 2), "%")
  )
) |>
  kbl(align = c("l","r")) |>
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)
```

## Daily Views & Reach

```{r daily-views-reach, fig.height=4.5}
growth_long <- growth |>
  select(Date, Views, Reach) |>
  pivot_longer(c(Views, Reach), names_to = "Metric", values_to = "Count")

ggplot(growth_long, aes(x = Date, y = Count, color = Metric, group = Metric)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.8, alpha = 0.7) +
  scale_color_manual(values = c("Views" = ucsb_blue, "Reach" = ucsb_gold)) +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "1 week", date_labels = "%b %d") +
  labs(
    title    = "Daily Profile Views & Reach — Spring Quarter 2026",
    x = NULL, y = "Count", color = NULL,
    caption  = "Source: UCSB Rec Instagram Profile Growth data"
  ) +
  theme_ucsb()
```

```{r views-reach-stats}
tibble(
  Metric        = c("Total Profile Views", "Total Reach", "Avg Daily Views", "Avg Daily Reach", "Peak Views Day"),
  Value = c(
    comma(sum(growth$Views, na.rm = TRUE)),
    comma(sum(growth$Reach,  na.rm = TRUE)),
    comma(round(mean(growth$Views, na.rm = TRUE))),
    comma(round(mean(growth$Reach,  na.rm = TRUE))),
    as.character(growth$Date[which.max(growth$Views)])
  )
) |>
  kbl(align = c("l","r")) |>
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)
```

---

# Post Performance {.tabset}

## Key Metrics

```{r post-kpis}
tibble(
  Metric  = c("Total Posts", "Avg Engagement Rate", "Avg Reach",
              "Avg Views", "Total Likes", "Total Saves", "Total Shares"),
  Value   = c(
    nrow(posts),
    paste0(round(mean(posts$Engagement_Rate_Pct), 2), "%"),
    comma(round(mean(posts$Reach))),
    comma(round(mean(posts$Views))),
    comma(sum(posts$Likes, na.rm = TRUE)),
    comma(sum(posts$Saves, na.rm = TRUE)),
    comma(sum(posts$Shares, na.rm = TRUE))
  )
) |>
  kbl(align = c("l","r")) |>
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)
```

## Engagement Rate by Media Type

```{r media-type-plot}
p_media <- posts |>
  ggplot(aes(x = Media_Type, y = Engagement_Rate_Pct, fill = Media_Type)) +
  geom_violin(alpha = 0.35, color = "white", trim = FALSE) +
  geom_boxplot(width = 0.15, outlier.shape = 21, outlier.size = 2.5,
               outlier.fill = ucsb_gold) +
  stat_summary(fun = mean, geom = "point", shape = 23,
               size = 3, fill = ucsb_gold, color = ucsb_navy) +
  scale_fill_manual(values = c("Carousel" = ucsb_blue, "Video" = ucsb_light)) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title    = "Engagement Rate by Media Type",
    subtitle = "Diamond = mean  |  Box = median & IQR  |  Spring Quarter 2026",
    x = NULL, y = "Engagement Rate",
    caption  = "Source: UCSB Rec Instagram Post data"
  ) +
  theme_ucsb() +
  theme(legend.position = "none")

p_media_reach <- posts |>
  ggplot(aes(x = Media_Type, y = Reach, fill = Media_Type)) +
  geom_boxplot(width = 0.4, outlier.shape = 21, outlier.size = 2.5,
               outlier.fill = ucsb_gold, alpha = 0.7) +
  scale_fill_manual(values = c("Carousel" = ucsb_blue, "Video" = ucsb_light)) +
  scale_y_continuous(labels = comma) +
  labs(
    title    = "Reach by Media Type",
    x = NULL, y = "Reach",
    caption  = " "
  ) +
  theme_ucsb() +
  theme(legend.position = "none")

p_media + p_media_reach
```

```{r media-summary-table}
posts |>
  group_by(Media_Type) |>
  summarise(
    `# Posts`       = n(),
    `Avg Eng Rate`  = paste0(round(mean(Engagement_Rate_Pct), 2), "%"),
    `Avg Reach`     = comma(round(mean(Reach))),
    `Avg Views`     = comma(round(mean(Views))),
    `Avg Likes`     = round(mean(Likes, na.rm = TRUE)),
    `Avg Shares`    = round(mean(Shares, na.rm = TRUE))
  ) |>
  kbl(align = c("l","c","r","r","r","r","r")) |>
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)
```

## Weekly Trends

```{r weekly-trends}
weekly <- posts |>
  group_by(Week) |>
  summarise(
    Total_Reach       = sum(Reach),
    Total_Engagements = sum(Engagements),
    Avg_Eng_Rate      = mean(Engagement_Rate_Pct),
    n_posts           = n()
  )

ggplot(weekly, aes(x = Week)) +
  geom_col(aes(y = Total_Reach), fill = ucsb_light, alpha = 0.85, width = 0.6) +
  geom_line(aes(y = Total_Engagements * 15), color = ucsb_blue, linewidth = 1.2) +
  geom_point(aes(y = Total_Engagements * 15), color = ucsb_gold, size = 3.5) +
  geom_text(aes(y = Total_Engagements * 15, label = Total_Engagements),
            vjust = -0.8, size = 3, color = ucsb_navy, fontface = "bold") +
  scale_y_continuous(
    name     = "Total Reach",
    labels   = comma,
    sec.axis = sec_axis(~ . / 15, name = "Total Engagements", labels = comma)
  ) +
  scale_x_continuous(breaks = weekly$Week) +
  labs(
    title    = "Weekly Reach & Engagements — Spring Quarter 2026",
    subtitle = "Bars = Total Reach  |  Line + labels = Total Engagements",
    x = "Week of Year",
    caption  = "Source: UCSB Rec Instagram Post data"
  ) +
  theme_ucsb()
```

## Best Day to Post

```{r day-of-week}
day_summary <- posts |>
  group_by(DayOfWeek) |>
  summarise(
    n_posts       = n(),
    Avg_Reach     = mean(Reach),
    Avg_Eng_Rate  = mean(Engagement_Rate_Pct)
  ) |>
  filter(!is.na(DayOfWeek))

p_day_reach <- ggplot(day_summary, aes(x = DayOfWeek, y = Avg_Reach, fill = Avg_Reach)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = comma(round(Avg_Reach))),
            vjust = -0.4, size = 3, color = ucsb_navy) +
  scale_fill_gradient(low = ucsb_light, high = ucsb_blue) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Avg Reach by Day of Week", x = NULL, y = "Avg Reach") +
  theme_ucsb()

p_day_er <- ggplot(day_summary, aes(x = DayOfWeek, y = Avg_Eng_Rate, fill = Avg_Eng_Rate)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(round(Avg_Eng_Rate, 1), "%")),
            vjust = -0.4, size = 3, color = ucsb_navy) +
  scale_fill_gradient(low = ucsb_light, high = ucsb_blue) +
  scale_y_continuous(labels = function(x) paste0(x, "%"),
                     expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Avg Engagement Rate by Day of Week", x = NULL, y = "Avg Eng Rate") +
  theme_ucsb()

p_day_reach / p_day_er
```

## Top 10 Posts

```{r top10-posts-chart, fig.height=5}
top10 <- posts |>
  arrange(desc(Engagement_Rate_Pct)) |>
  head(10) |>
  mutate(Label = str_wrap(str_trunc(Caption_Short, 45), 38))

ggplot(top10, aes(x = Engagement_Rate_Pct,
                  y = reorder(Label, Engagement_Rate_Pct),
                  fill = Media_Type)) +
  geom_col(alpha = 0.85) +
  geom_text(aes(label = paste0(Engagement_Rate_Pct, "%")),
            hjust = -0.1, size = 3.2, color = ucsb_navy, fontface = "bold") +
  scale_fill_manual(values = c("Carousel" = ucsb_blue, "Video" = ucsb_light)) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.18)),
                     labels = function(x) paste0(x, "%")) +
  labs(
    title    = "Top 10 Posts by Engagement Rate — Spring Quarter 2026",
    x = "Engagement Rate", y = NULL, fill = "Media Type",
    caption  = "Source: UCSB Rec Instagram Post data"
  ) +
  theme_ucsb()
```

```{r top10-posts-table}
posts |>
  arrange(desc(Engagement_Rate_Pct)) |>
  head(10) |>
  select(Date, Caption_Short, Media_Type, Engagement_Rate_Pct, Reach, Likes, Saves, Shares) |>
  mutate(Engagement_Rate_Pct = paste0(Engagement_Rate_Pct, "%")) |>
  rename(
    `Date`      = Date,
    `Caption`   = Caption_Short,
    `Type`      = Media_Type,
    `Eng Rate`  = Engagement_Rate_Pct,
    `Reach`     = Reach,
    `Likes`     = Likes,
    `Saves`     = Saves,
    `Shares`    = Shares
  ) |>
  kbl() |>
  kable_styling(bootstrap_options = c("striped","hover","condensed")) |>
  column_spec(4, bold = TRUE, color = ucsb_blue)
```

---

# Reel Performance {.tabset}

## Key Metrics

```{r reel-kpis}
tibble(
  Metric  = c("Total Reels", "Avg Engagement Rate", "Avg Views",
              "Avg Reach", "Avg Skip Rate", "Total Shares"),
  Value   = c(
    nrow(reels),
    paste0(round(mean(reels$Engagement_Rate_Pct), 2), "%"),
    comma(round(mean(reels$Views))),
    comma(round(mean(reels$Reach))),
    paste0(round(mean(reels$Skip_Rate_Pct, na.rm = TRUE), 1), "%"),
    comma(sum(reels$Shares))
  )
) |>
  kbl(align = c("l","r")) |>
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)
```

## Views & Engagement Over Time

```{r reel-trends}
p_reel_views <- reels |>
  arrange(Time_Posted) |>
  ggplot(aes(x = Date, y = Views)) +
  geom_col(fill = ucsb_blue, alpha = 0.8, width = 2) +
  geom_text(aes(label = comma(Views)), vjust = -0.4, size = 2.8, color = ucsb_navy) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
  scale_x_date(date_breaks = "1 week", date_labels = "%b %d") +
  labs(title = "Reel Views Over Time", x = NULL, y = "Views") +
  theme_ucsb()

p_reel_er <- reels |>
  arrange(Time_Posted) |>
  ggplot(aes(x = Date, y = Engagement_Rate_Pct)) +
  geom_line(color = ucsb_gold, linewidth = 1.1) +
  geom_point(color = ucsb_blue, size = 3) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_x_date(date_breaks = "1 week", date_labels = "%b %d") +
  labs(title = "Reel Engagement Rate Over Time", x = NULL, y = "Engagement Rate") +
  theme_ucsb()

p_reel_views / p_reel_er
```

## Skip Rate Analysis

```{r skip-rate, fig.height=4}
reels |>
  filter(!is.na(Skip_Rate_Pct)) |>
  ggplot(aes(x = Skip_Rate_Pct, y = Engagement_Rate_Pct)) +
  geom_smooth(method = "lm", se = TRUE,
              color = ucsb_light, fill = ucsb_grey, alpha = 0.3, linewidth = 0.8) +
  geom_point(aes(size = Views, color = Views), alpha = 0.85) +
  scale_color_gradient(low = ucsb_gold, high = ucsb_blue, labels = comma) +
  scale_size_continuous(range = c(4, 12), labels = comma) +
  scale_x_continuous(labels = function(x) paste0(x, "%")) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title    = "Reel Skip Rate vs. Engagement Rate",
    subtitle = "Bubble size and color = Views. Lower skip + higher engagement = best reels.",
    x = "Skip Rate", y = "Engagement Rate",
    color = "Views", size = "Views",
    caption  = "Source: UCSB Rec Instagram Reel data"
  ) +
  theme_ucsb()
```

## Top Reels

```{r top-reels-table}
reels |>
  arrange(desc(Views)) |>
  select(Date, Caption_Short, Engagement_Rate_Pct, Views, Reach, Likes, Shares, Skip_Rate_Pct) |>
  mutate(
    Engagement_Rate_Pct = paste0(Engagement_Rate_Pct, "%"),
    Skip_Rate_Pct       = ifelse(is.na(Skip_Rate_Pct), "—", paste0(Skip_Rate_Pct, "%"))
  ) |>
  rename(
    `Caption`    = Caption_Short,
    `Eng Rate`   = Engagement_Rate_Pct,
    `Skip Rate`  = Skip_Rate_Pct
  ) |>
  kbl() |>
  kable_styling(bootstrap_options = c("striped","hover","condensed")) |>
  column_spec(3, bold = TRUE, color = ucsb_blue)
```

---

# Key Takeaways

```{r takeaways-setup, include=FALSE}
top_post        <- posts |> arrange(desc(Engagement_Rate_Pct)) |> slice(1)
top_reel        <- reels |> arrange(desc(Views)) |> slice(1)
best_day        <- posts |> group_by(DayOfWeek) |>
                   summarise(avg = mean(Reach)) |> arrange(desc(avg)) |> slice(1)
avg_er_spring   <- round(mean(posts$Engagement_Rate_Pct), 2)
avg_er_winter   <- 3.65   # from Winter Quarter report
er_change       <- round(avg_er_spring - avg_er_winter, 2)
```

- **Engagement rate improved** from Winter (`r avg_er_winter`%) to Spring (`r avg_er_spring`%) — a **+`r er_change` percentage point increase**.
- **Carousels remain the top-performing format** with higher average engagement rates and reach than videos.
- **No image-only posts** were published this quarter — a shift from Winter where 5 image posts were included.
- **Top post:** "`r str_trunc(top_post$Caption_Short, 60)`" with `r top_post$Engagement_Rate_Pct`% engagement rate.
- **Top reel by views:** "`r str_trunc(top_reel$Caption_Short, 60)`" with `r comma(top_reel$Views)` views.
- **Best posting day:** `r best_day$DayOfWeek` by average reach.
- **Follower growth:** +`r net_growth` followers over the quarter (`r follower_start` → `r follower_end`).
- **Profile views:** `r comma(sum(growth$Views))` total profile views across the quarter.

---

